Volatility and the Buyback Anomaly: Interactive Tool

### 2017-02-17 manuscript.Rnw
clean up

<< eval = TRUE, echo=FALSE,message=FALSE,fig.pos='h',results='asis' >>=
#dev.off();
#library(knitr); knit2pdf('manuscriptRF.Rnw', texi2dvi='texi2dvi'); file.remove(paste("manuscriptRF",c(".aux",".tex",".blg",".bbl",".log",".out"),sep=""))
#dir.create("tmpfiles") # need to comment this if the directory already exists
# Then to make it into Word: 
# system("pandoc -s manuscriptRF.tex --bibliography=bibliography.bib --number-sections --reference-links -o theo.docx")

@@

<< fig.width=10, fig.height=6, out.width='\\textwidth', echo=F >>=
months1 = cumsum(100*remove_initialization_time(BUYBACK_DATA$long_all1mshort_risk_factors,min_date=FirstTrade))
months3 = cumsum(100*remove_initialization_time(BUYBACK_DATA$long_all3mshort_risk_factors,min_date=FirstTrade))
months6 = cumsum(100*remove_initialization_time(BUYBACK_DATA$long_all6mshort_risk_factors,min_date=FirstTrade))
months12 = cumsum(100*remove_initialization_time(BUYBACK_DATA$long_all12mshort_risk_factors,min_date=FirstTrade))
months24 = cumsum(100*remove_initialization_time(BUYBACK_DATA$long_all24mshort_risk_factors,min_date=FirstTrade))
months36 = cumsum(100*remove_initialization_time(BUYBACK_DATA$long_all36mshort_risk_factors,min_date=FirstTrade))
months48 = cumsum(100*remove_initialization_time(BUYBACK_DATA$long_all48mshort_risk_factors,min_date=FirstTrade))

@@

So far all our events are equally weighted. \cite{MitchellStafford2000} argue that events should be value weighted to test whether they represent an economically important anomaly. However, as pointed out by \cite{LoughranRitter2000}, small firms are more likely to be mispriced as they are less followed by analysts. Hence managers of small firms are better able to take advantage of mispricing than managers of large firms, which is the reason why firm size is one of the components of the U-index. So value weighting the events would simply bias the results toward zero. And indeed, when we value-weight the events (see Table \ref{tbl:valuetable} Panel A) long-term excess returns become statistically insignificantly different from zero when using the total sample of events. So the buyback anomaly is not economically important and does not challenge the basic premise that “the market” represented by a value-weighted index is priced correctly. Note, however, if we eliminate the firms with are in the top $25^{th}$ market capitalization percentile (see Table \ref{tbl:valuetable}, Panel~B) the results become again significant, especially for the high U-index firms. These firms had a market capitalization less than \$2,971 million in 2015. The fact that the buyback anomaly is a small cap anomaly makes it more likely that excess returns are evidence that managers (at least in these firms) time the market based on their information advantage relative to investors.  It may also explain why the anomaly persists as the capacity of small cap funds (and therefore management fees) is smaller than the capacity of large cap funds.

@@

<<  eval = TRUE, echo=FALSE,message=FALSE,fig.pos='h',results='asis' >>=
#EU_index_features = scrub(cbind(BUYBACK_DATA$DATASET$SDC$minus_Rsq_returns, BUYBACK_DATA$DATASET$SDC$pre_vol,BUYBACK_DATA$Valuation_Index))
#colnames(EU_index_features) <- c("Idiosyncratic Score", "Volatility Score", "U-Index Score")
#Under_IdioBB_cor = cor(EU_index_features)
@

@@

<<   eval = TRUE, echo=FALSE,message=FALSE,fig.pos='h',results='asis' >>=
printLatexTable2(
  cbind(IRATS_table_all_iss,IRATS_table_lowMM_iss,IRATS_table_highMM_iss)[reported_times,],
  cbind(CAL_table_all_iss,CAL_table_lowvol_iss,CAL_table_highvol_iss)[reported_times,],
  columns = c("All 4F","All 5F","Low MM 4F","Low MM 5F","High MM 4F","High MM 5F"),
  title="SEO announcements during 1985-2015",
  title1 = "Panel A: IRATs Cumulative Abnormal Returns",
  title2 = "Panel B: Calendar Method Monthly Abnormal Returns",
  caption= "The table presents the abnormal returns for firms after issue announcements from the announcement date until $t$ months after the announcement. We include a version of the abnormal returns for the full sample and one for both companies with high low Mispricing Measure (MM) before the announcement.  Panel~A reports monthly cumulative average abnormal returns (CAR) in percent using \\cite{Ibbotson1975} returns across time and security (IRATS) method combined with the \\cite{StambaughYuan2016} four-factor model (4F) and the \\cite{FamaFrench2015} five-factor model (5F) for the sample of firms that announced equity issuance plus various subsamples. The following regressions are run each event month~$j$:\\vspace{-1em}
\\begin{eqnarray*}
(R_{i,t} - R_{f,t}) &=& a_j + b_j (R_{m,t} - R_{f,t}) + c_j {SMB'}_t + d_j {MGMT}_t + e_t {PERF}_t + \\epsilon_{i,t}, \\\\
(R_{i,t} - R_{f,t}) &=& a_j + b_j (R_{m,t} - R_{f,t}) + c_j {SMB}_t + d_j {HML}_t + e_t {RMW}_t + f_t {CMA}_t + \\epsilon_{i,t},
\\end{eqnarray*}
where $R_{i,t}$ is the monthly return on security $i$ in the calendar month $t$ that corresponds to the event month $j$, with $j = 0$ being the month of the equity issuance. $R_{f,t}$ and $R_{m,t}$ are the risk-free rate and the return on the equally weighted CRSP index, respectively. ${SMB}_t$, ${HML}_t$, ${RMW}_t$, ${CMA}_t$ are the monthly returns on the size, book-to-market factor, profitability factor and investment factor in month $t$, respectively. ${SMB'}_t$, ${MGMT}_t$, ${PERF}_t$ are the \\cite{StambaughYuan2016} factors Size, Management Related (MGMT), and  Performance Related (PERF), respectively. The numbers reported are sums of the intercepts of cross-sectional regressions over the relevant event-time-periods expressed in percentage terms. The standard error (denominator of the $t$-statistic) for a window is the square root of the sum of the squares of the monthly standard errors. Panel~B reports monthly average abnormal returns (AR) of equally weighted Calendar Time portfolios. In this method, event firms that have announced an equity issuance in the last calendar months form the basis of the calendar month portfolio. A single time-series regression is run with the excess returns of the calendar portfolio as the dependent variable and the returns of the \\cite{StambaughYuan2016} four factors (4F) or the \\cite{FamaFrench2015} five factors (5F) as the independent variables. The significance levels are indicated by +, *, and ** and correspond to a significance level of $10\\%$, $5\\%$, and $1\\%$ respectively, using a two-tailed test.
",

@@

<< fig.width=8, fig.height=5, out.width='\\textwidth', echo=F >>=
R2_IRATStableBB_plot = R2_IRATStableBB
R2_IRATStableBB_plot[rownames(R2_IRATStableBB_plot) %in% as.character((-6):(-1)),c("Low Idiosync.: CAR","High Idiosync.: CAR")] <- -R2_IRATStableBB_plot[rownames(R2_IRATStableBB_plot) %in% as.character((-6):(-1)),c("Low Idiosync.: CAR","High Idiosync.: CAR")] 
R2_IRATStable_underBB_plot = R2_IRATStable_underBB
R2_IRATStable_underBB_plot[rownames(R2_IRATStable_underBB_plot) %in% as.character((-6):(-1)),c(1,4,7,10)] <- -R2_IRATStable_underBB_plot[rownames(R2_IRATStable_underBB_plot) %in% as.character((-6):(-1)),c(1,4,7,10)] 
@

<< fig.width=8, fig.height=5, out.width='\\textwidth', echo=F >>=

VOL_IRATStableBB_plot = VOL_IRATStableBB
VOL_IRATStableBB_plot[rownames(VOL_IRATStableBB_plot) %in% as.character((-6):(-1)),c("Low Vol: CAR","High Vol: CAR")] <- -VOL_IRATStableBB_plot[rownames(VOL_IRATStableBB_plot) %in% as.character((-6):(-1)),c("Low Vol: CAR","High Vol: CAR")] 
VOL_IRATStable_underBB_plot = VOL_IRATStable_underBB
VOL_IRATStable_underBB_plot[rownames(VOL_IRATStable_underBB_plot) %in% as.character((-6):(-1)),c(1,4,7,10)] <- -VOL_IRATStable_underBB_plot[rownames(VOL_IRATStable_underBB_plot) %in% as.character((-6):(-1)),c(1,4,7,10)] 
@

>>>>

<< eval = TRUE, echo=FALSE,message=FALSE,fig.pos='h',results='asis' >>=
#dev.off();
#library(knitr); knit2pdf('manuscript.Rnw', texi2dvi='texi2dvi'); file.remove(paste("manuscript",c(".aux",".tex",".blg",".bbl",".log",".out"),sep=""))
#dir.create("tmpfiles") # need to comment this if the directory already exists
# Then to make it into Word: 
# system("pandoc -s manuscript.tex --bibliography=bibliography.bib --number-sections --reference-links -o theo.docx")

@@

@@

So far all our events are equally weighted. \cite{MitchellStafford2000} argue that events should be value weighted to test whether they represent an economically important anomaly. However, as pointed out by \cite{LoughranRitter2000}, small firms are more likely to be mispriced as they are less followed by analysts. Hence managers of small firms are better able to take advantage of mispricing than managers of large firms, which is the reason why firm size is one of the components of the U-index. So value weighting the events would simply bias the results toward zero. And indeed, when we value-weight the events (see Table \ref{tbl:valuetable} Panel A) long-term excess returns become statistically insignificantly different from zero when using the total sample of events. So the buyback anomaly is not economically important and does not challenge the basic premise that ???the market??? represented by a value-weighted index is priced correctly. Note, however, if we eliminate the firms with are in the top $25^{th}$ market capitalization percentile (see Table \ref{tbl:valuetable}, Panel~B) the results become again significant, especially for the high U-index firms. These firms had a market capitalization less than \$2,971 million in 2015. The fact that the buyback anomaly is a small cap anomaly makes it more likely that excess returns are evidence that managers (at least in these firms) time the market based on their information advantage relative to investors.  It may also explain why the anomaly persists as the capacity of small cap funds (and therefore management fees) is smaller than the capacity of large cap funds.

@@

@@

<<   eval = TRUE, echo=FALSE,message=FALSE,fig.pos='h',results='asis' >>=
printLatexTable2(
  cbind(IRATS_table_all_iss,IRATS_table_lowMM_iss,IRATS_table_highMM_iss)[reported_times,],
  cbind(CAL_table_all_iss,CAL_table_lowMM_iss,CAL_table_highMM_iss)[reported_times,],
  columns = c("All 4F","All 5F","Low MM 4F","Low MM 5F","High MM 4F","High MM 5F"),
  title="SEO announcements during 1985-2015",
  title1 = "Panel A: IRATs Cumulative Abnormal Returns",
  title2 = "Panel B: Calendar Method Monthly Abnormal Returns",
  caption= "The table presents the abnormal returns for firms after issue announcements from the announcement date until $t$ months after the announcement. We include a version of the abnormal returns for the full sample and one for both companies with high low Mispricing Measure (MM) before the announcement.  Panel~A reports monthly cumulative average abnormal returns (CAR) in percent using \\cite{Ibbotson1975} returns across time and security (IRATS) method combined with the \\cite{StambaughYuan2016} four-factor model (4F) and the \\cite{FamaFrench2015} five-factor model (5F) for the sample of firms that announced equity issuance plus various subsamples. The following regressions are run each event month~$j$:\\vspace{-1em}
\\begin{eqnarray*}
(R_{i,t} - R_{f,t}) &=& a_j + b_j (R_{m,t} - R_{f,t}) + c_j {SMB'}_t + d_j {MGMT}_t + e_t {PERF}_t + \\epsilon_{i,t}, \\\\
(R_{i,t} - R_{f,t}) &=& a_j + b_j (R_{m,t} - R_{f,t}) + c_j {SMB}_t + d_j {HML}_t + e_t {RMW}_t + f_t {CMA}_t + \\epsilon_{i,t},
\\end{eqnarray*}
where $R_{i,t}$ is the monthly return on security $i$ in the calendar month $t$ that corresponds to the event month $j$, with $j = 0$ being the month of the equity issuance. $R_{f,t}$ and $R_{m,t}$ are the risk-free rate and the return on the equally weighted CRSP index, respectively. ${SMB}_t$, ${HML}_t$, ${RMW}_t$, ${CMA}_t$ are the monthly returns on the size, book-to-market factor, profitability factor and investment factor in month $t$, respectively. ${SMB'}_t$, ${MGMT}_t$, ${PERF}_t$ are the \\cite{StambaughYuan2016} factors Size, Management Related (MGMT), and  Performance Related (PERF), respectively. The numbers reported are sums of the intercepts of cross-sectional regressions over the relevant event-time-periods expressed in percentage terms. The standard error (denominator of the $t$-statistic) for a window is the square root of the sum of the squares of the monthly standard errors. Panel~B reports monthly average abnormal returns (AR) of equally weighted Calendar Time portfolios. In this method, event firms that have announced an equity issuance in the last calendar months form the basis of the calendar month portfolio. A single time-series regression is run with the excess returns of the calendar portfolio as the dependent variable and the returns of the \\cite{StambaughYuan2016} four factors (4F) or the \\cite{FamaFrench2015} five factors (5F) as the independent variables. The significance levels are indicated by +, *, and ** and correspond to a significance level of $10\\%$, $5\\%$, and $1\\%$ respectively, using a two-tailed test.
",

@@
